---
title: "Data Prep: AUSTRALIA cars"
---

## OPTION 1 ##
# AUSTRALIA CAR REGISTRATIONS (From gov.) + ELECTRIC CAR SALES (From Electric Vehicle Council)
Read the "AUbefore2015" excel file and separate the column out.
```{r}
library(readxl)
library(tidyverse)

columns <- c("Year","LeadedPetrol","UnleadedPetrol","TotalPetrol","Diesel",
             "Other(InclLPG+DualFuel+Electric)","Total")
AUbefore2015 <- read_excel("AUbefore2015.xlsx",sheet="Data")
AUbefore2015 %>% extract(col=1,
                         into=columns, regex="(20[0-9]{2})(\\s[0-9\\s]{0,3}[0-9]{3}\\s[0-9]{3})(\\s[0-9\\s]{0,3}[0-9]{3}\\s[0-9]{3})(\\s[0-9\\s]{0,3}[0-9]{3}\\s[0-9]{3})(\\s[0-9\\s]{0,3}[0-9]{3}\\s[0-9]{3})(\\s[0-9\\s]{0,3}[0-9]{3}\\s[0-9]{3})(\\s[0-9\\s]{0,3}[0-9]{3}\\s[0-9]{3})") %>% 
  mutate_each(funs(if(is.character(.)) gsub(" ","",.) else .)) %>% 
  mutate_each(funs(if(!grepl("20[0-9]{2}",.)) as.numeric(.) else .)) -> AUbaseFuels_untidy
AUbaseFuels_untidy

```


From each Excel file for 2015 and after, extract the data in the tab "Table_5" on line 43. Make to sure to run the earlier code block BEFORE running this each time.
```{r}
# Read from table 5, line 43 and add the following column headers.
lines2skip <- 42
sheet_name <- "Table_5"

file_names <- c("AU2015.xls","AU2016.xls","AU2017.xls","AU2018.xls","AU2019.xls","AU2020.xls")
for(i in 1:length(file_names)) {
  temp <- read_excel(file_names[i],sheet=sheet_name, col_names=columns, skip=lines2skip)
  temp %>% head(1) -> temp
  AUbaseFuels_untidy <- rbind(temp,AUbaseFuels_untidy)
  rm(temp)
}
AUbaseFuels_untidy %>%
  select(-c(LeadedPetrol,UnleadedPetrol)) %>% 
  rename(Petrol=TotalPetrol) -> AUbaseFuels
AUbaseFuels
```

Now add the data from the June 2018 "The state of electric vehicles in Australia" report. Note that this is SALES data rather than REGISTRATION data. But it's the best I could find.
```{r}
electric <- read_excel("AU_ElectricCarSales2018.xlsx",sheet="Data")
electric %>% mutate(Year=as.character(Year)) -> electric
electric
AUbaseFuels %>%
  left_join(electric, by=c("Year"="Year")) %>% 
  mutate(Other=`Other(InclLPG+DualFuel+Electric)`-ifelse(is.na(Electric),0,Electric),
         Hybrid=NA,
         "LPG/NaturalGas"=NA,
         Region="Australia") %>% 
  select(Region, Year, Petrol, Diesel, Hybrid, Electric, `LPG/NaturalGas`, Other) -> AUcars1
AUcars1
```




## OPTION 2 ##
# FROM NTC (National Transport Commission)
```{r}
AUcars2_untidy <- read_excel("Australia_CO2_reports.xlsx",sheet="Data")
AUcars2_untidy %>%
  mutate(Region="Australia",
         "LPG/NaturalGas"=LPG,
         Other=NA,
         Hybrid=NA,
         Year=as.character(Year),
         Electric=`Electric (Incl Tesla)`) %>% 
  select(Region, Year, Petrol, Diesel, Hybrid, Electric, `LPG/NaturalGas`, Other) -> AUcars2
AUcars2
```




# EXPORT (Remember to select which dataframe you want)
```{r}
# Export to excel file so it can be easily processed after
library(xlsx)
write.xlsx(AUcars2, "AUSTRALIACarCountByFuel.xlsx")
```

