---
title: "Data Prep: BRAZIL Cars"
---

Read in the data.
```{r}
library(tidyverse)
library(readxl)
BrazilCars_untidy <- read_excel("Brazil-LightVehicleRegis-2014-2019-by-fuel.xlsx", sheet="Data", skip=4)
BrazilCars_untidy
```


Reorganize it so it matches the same format as the other countries.
```{r}
BrazilCars_untidy %>% 
  mutate(Region="Brazil",
         Year=`...1`,
         Petrol=Gasoline,
         Hybrid=`Flexible fuel*`,
         Electric=`Electric**`,
         "LPG/NaturalGas"=NA,
         Other=Ethanol) %>% 
  select(Region, Year, Petrol, Diesel, Hybrid, Electric, `LPG/NaturalGas`, Other) -> BrazilCars
BrazilCars
```
ISSUE: The "Hybrid" section here actually contains vehicles that run on Petrol/Ethanol and is NOT consistent with my definition for hybrid electric vehicles. Hybrid electric vehicles are incorporated in the "Electric" category for Brazil.
CONCLUSION: It might make more sense to add the Flexible Fuel cars to "Petrol" since they use Petrol/Ethanol (Ethanol is categorized as an "Other" fuel). The "Hybrid" category would then be empty (NULL).

```{r}
BrazilCars %>% 
  mutate(Petrol=Petrol+Hybrid, Hybrid=NA) -> BrazilCars2
BrazilCars2
```


```{r}
# Export to excel file so it can be easily processed after
library(xlsx)
write.xlsx(BrazilCars2, "BRAZILcars.xlsx")
```


