---
title: "Data Prep: EUROPE Cars"
---
## IN THIS VERSION, ALL COUNTRIES ARE KEPT

```{r}
# Read in data from the excel file
library(tidyverse)
library(readxl)
xlfile <- "EUROPEFuelType-CarCount.xlsx"
CarCount2019_untidy <- read_excel(xlfile, sheet="Count_2019Report")
CarCount2018_untidy <- read_excel(xlfile, sheet="Count_2018Report")
CarCount2017_untidy <- read_excel(xlfile, sheet="Count_2017Report")
FuelType2019.1_untidy <- read_excel(xlfile, sheet="Fuels1_2019Report")
FuelType2019.2_untidy <- read_excel(xlfile, sheet="Fuels2_2019Report")
FuelType2018.1_untidy <- read_excel(xlfile, sheet="Fuels1_2018Report")
FuelType2018.2_untidy <- read_excel(xlfile, sheet="Fuels2_2018Report")
FuelType2018.3_untidy <- read_excel(xlfile, sheet="Fuels3_2018Report")
FuelType2017.1_untidy <- read_excel(xlfile, sheet="Fuels1_2017Report")
FuelType2017.2_untidy <- read_excel(xlfile, sheet="Fuels2_2017Report")
```

This contains data on:

* "ACEA Report Vehicles in use Europe 2019"
* "ACEA Report Vehicles in use Europe 2018"
* "ACEA Report Vehicles in use Europe 2017"


```{r}
# View of 2019 car count data (untidied)
CarCount2019_untidy

# Cleaning the 2019 car count data
CarCount2019_untidy %>%
  extract(`20142015201620172018%change18/17`, c("Region",2014:2018,"%change18/17"), 
          "([A-Za-z\\s]+)([0-9]{0,3}[,]?[0-9]{0,3},[0-9]{0,3})([0-9]{0,3}[,]?[0-9]{0,3},[0-9]{0,3})([0-9]{0,3}[,]?[0-9]{0,3},[0-9]{0,3})([0-9]{0,3}[,]?[0-9]{0,3},[0-9]{0,3})([0-9]{0,3}[,]?[0-9]{0,3},[0-9]{0,3})(.*)") %>% 
  pivot_longer(cols=as.character(c(2014:2018)),names_to="Year",values_to="Count1") %>%
  mutate(Count=as.numeric(gsub(",", "", `Count1`))) %>% 
  select(`Report Year`, `Region`, `Vehicle Type`, `Year`, `Count`) %>% 
  arrange(`Region`, `Vehicle Type`, desc(`Year`)) -> CarCount2019
CarCount2019
```

```{r}
# The 2018 and 2017 car count data is about the same.
clean_it <- function(x) {
  x %>%
    pivot_longer(cols=contains("0"),names_to="Year",values_to="Count1") %>% 
    mutate(Count=as.numeric(`Count1`)) %>% 
    select(`Report Year`, `Region`, `Vehicle Type`, `Year`, `Count`) %>% 
    arrange(`Region`, `Vehicle Type`, desc(`Year`)) -> x_clean
  x_clean
}
(CarCount2018 <- clean_it(CarCount2018_untidy))
(CarCount2017 <- clean_it(CarCount2017_untidy))
```

A display of the tidied 2018 and 2017 car count data.


```{r}
# Merge the car count data from each year together where the latest report has the highest priority.
CarCount2019 %>% 
  rbind(CarCount2018, CarCount2017) %>% 
  unite(id,Region,`Vehicle Type`,Year,sep="_") %>%
  filter(!duplicated(id)) %>% # Keep data from latest report
  filter(!grepl("[A-Za-z\\s]+_Total commercial vehicles \\(incl buses\\)",id)) %>% # Removed repeated data
  filter(!grepl(".*\\(incl buses\\)_201[2-5]",id)) %>% # Removed repeated data
  separate(id,into=c("Region","Vehicle Type","Year"),sep="_") %>% 
  arrange(`Region`, `Vehicle Type`, desc(`Year`)) -> CarCountALL
CarCountALL %>% unite(id,Region,`Vehicle Type`,Year,sep="_") -> CarCountALL_id
CarCountALL_id
```

In this dataset, the car count data from each report was merged. The latest records were kept for each year and repeated data was removed.


```{r}
# The fuel data for some of the 2018 and 2017 files look the same. Use a function for it.
clean_it2 <- function(x) {
  x %>% 
    mutate("Electric(BatteryElectric+Plug-inHybrid)"=`Electric(inclPlugIn)`) %>% 
    select(-`Electric(inclPlugIn)`) %>% 
    pivot_longer(Petrol:"Electric(BatteryElectric+Plug-inHybrid)",names_to="Fuel Type",values_to="% Share") %>% 
    select(`Report Year`,`Region`,`Vehicle Type`, `Fuel Type`, `Year`, `% Share`) -> x_clean
  x_clean
}
FuelType2017.1 <- clean_it2(FuelType2017.1_untidy)
FuelType2018.2 <- clean_it2(FuelType2018.2_untidy)

clean_it3 <- function(x) {
  x %>% 
    mutate("Electric(BatteryElectric+Plug-inHybrid)"=`Electric`) %>% 
    select(-`Electric`) %>% 
    pivot_longer(Petrol:"Electric(BatteryElectric+Plug-inHybrid)",names_to="Fuel Type",values_to="% Share") %>% 
    select(`Report Year`,`Region`,`Vehicle Type`, `Fuel Type`, `Year`, `% Share`) -> x_clean
  x_clean
}
FuelType2018.3 <- clean_it3(FuelType2018.3_untidy)
FuelType2017.2 <- clean_it3(FuelType2017.2_untidy)

FuelType2018.1_untidy %>%
  mutate("Electric(BatteryElectric+Plug-inHybrid)"=`Electric(InclFuelCell)`+`Plug‐inHybrids`) %>% 
  select(-c(`Electric(InclFuelCell)`,`Plug‐inHybrids`)) %>% 
  pivot_longer(Petrol:"Electric(BatteryElectric+Plug-inHybrid)",names_to="Fuel Type",values_to="% Share") %>% 
  select(`Report Year`,`Region`,`Vehicle Type`, `Fuel Type`, `Year`, `% Share`) -> FuelType2018.1

FuelTypeALL <- rbind(FuelType2018.1,FuelType2018.2,FuelType2018.3,FuelType2017.1,FuelType2017.2)
FuelTypeALL %>%
  unite(id,Region,`Vehicle Type`,Year,sep="_") %>% 
  arrange(id) -> FuelTypeALL_id
FuelTypeALL_id
```

In this dataset, the fuel data from the 2017 and 2018 report was merged.


```{r}
# View of 2019 fuel type data (untidied)
FuelType2019.1_untidy
FuelType2019.2_untidy

# Fix up the 2019 fuel type data for passenger cars
FuelType2019.1_untidy %>% 
  extract(`Petrol Diesel HybridElectric BatteryElectric Plug-inHybrid LPG+NaturalGas Other+Unknown`,
          c("Region","Petrol1","Diesel1","HybridElectric1","BatteryElectric1","Plug-inHybrid1","LPG/NaturalGas1","Other1"),
          "([A-Za-z\\s]+)([0-9.]+%)([0-9.]+%)([0-9.]+%)([0-9.]+%)([0-9.]+%)([0-9.]+%)([0-9.]+%)") %>%
  mutate("Petrol"=as.numeric(gsub("%", "", `Petrol1`))/100,
         "Diesel"=as.numeric(gsub("%", "", `Diesel1`))/100,
         "Hybrids"=as.numeric(gsub("%", "", `HybridElectric1`))/100,
         "BatteryElectric"=as.numeric(gsub("%", "", `BatteryElectric1`))/100,
         "Plug-inHybrid"=as.numeric(gsub("%", "", `Plug-inHybrid1`))/100,
         "LPG/NaturalGas"=as.numeric(gsub("%", "", `LPG/NaturalGas1`))/100,
         "Other"=as.numeric(gsub("%", "", `Other1`))/100) %>%
  mutate("Electric(BatteryElectric+Plug-inHybrid)"=`BatteryElectric`+`Plug-inHybrid`) %>% 
  select(-c(`Metric`,`BatteryElectric`,`Plug-inHybrid`),-contains("1")) %>% 
  pivot_longer(cols = c("Petrol","Diesel","Hybrids","Electric(BatteryElectric+Plug-inHybrid)","LPG/NaturalGas","Other"),
               names_to = "Fuel Type", values_to = "% Share") %>% 
  unite(id,Region,`Vehicle Type`,Year,sep="_") -> FuelType2019.1
FuelType2019.1

# Fix up the rest of the 2019 fuel type data
FuelType2019.2_untidy %>% 
  extract(`Petrol Diesel HybridElectric Electric(BatteryElectric+Plug-inHybrid) LPG+NaturalGas Other+Unknown`,
          c("Region","Petrol","Diesel","Hybrids","Electric(BatteryElectric+Plug-inHybrid)","LPG/NaturalGas","Other"),
          "([A-Za-z\\s]+)([0-9.]+%)([0-9.]+%)([0-9.]+%)([0-9.]+%)([0-9.]+%)([0-9.]+%)") %>% 
  pivot_longer(cols = c("Petrol","Diesel","Hybrids","Electric(BatteryElectric+Plug-inHybrid)","LPG/NaturalGas","Other"),
               names_to = "Fuel Type", values_to = "% Share 1") %>% 
  mutate("% Share"=as.numeric(gsub("%", "", `% Share 1`))/100) %>%
  select(-c(`Metric`,`% Share 1`)) %>% 
  unite(id,Region,`Vehicle Type`,Year,sep="_") -> FuelType2019.2
FuelType2019.2

# Merge with the remaining fuel type data
FuelTypeALL_id %>% 
  rbind(FuelType2019.1,FuelType2019.2) %>% 
  arrange(id) -> FuelTypeALL_id
FuelTypeALL_id
```
In this dataset, the fuel data from the 2019 report was cleaned.


```{r}
# Create a dataset containing the car count based on fuel type
FuelTypeALL_id %>%
  left_join(CarCountALL_id, by=c("id"="id")) %>% 
  mutate("Num Vehicles for Fuel Type"=`% Share`*`Count`) %>% 
  mutate("% Share for Fuel Type"=`% Share`) %>% 
  mutate("Total Vehicles (All Fuel Types)"=`Count`) %>% 
  separate(id,into=c("Region","Vehicle Type","Year"),sep="_") %>%
  select(`Region`,`Vehicle Type`, `Fuel Type`, `Year`, `% Share for Fuel Type`,
         `Num Vehicles for Fuel Type`, `Total Vehicles (All Fuel Types)`) %>% 
  arrange(`Region`,`Vehicle Type`, `Fuel Type`, `Year`) -> CarCountByFuel
CarCountByFuel[CarCountByFuel=="United Kingdom"]<-""
CarCountByFuel
```


```{r}
# Export to excel file so it can be easily processed after
library(xlsx)
write.xlsx(CarCountByFuel, "EUROPEcars.xlsx")
```

