---
title: "Data Prep: NORWAY-NETHERLANDS Cars"
---
## THIS IS OUTDATED AND WILL NOT WORK WITH THE NEW EXCEL FILE

```{r}
# Read in data from the excel file
library(tidyverse)
library(readxl)
xlfile <- "EUROPEFuelType-CarCount.xlsx"
CarCount2019_untidy <- read_excel(xlfile, sheet="Count_2019Report")
CarCount2018_untidy <- read_excel(xlfile, sheet="Count_2018Report")
CarCount2017_untidy <- read_excel(xlfile, sheet="Count_2017Report")
FuelType2019 <- read_excel(xlfile, sheet="Fuels_2019Report_Raw")
FuelType2018.1_untidy <- read_excel(xlfile, sheet="Fuels1_2018Report")
FuelType2018.2_untidy <- read_excel(xlfile, sheet="Fuels2_2018Report")
FuelType2018.3_untidy <- read_excel(xlfile, sheet="Fuels3_2018Report")
FuelType2017.1_untidy <- read_excel(xlfile, sheet="Fuels1_2017Report")
FuelType2017.2_untidy <- read_excel(xlfile, sheet="Fuels2_2017Report")
```

This contains data on:

* "ACEA Report Vehicles in use Europe 2019"
* "ACEA Report Vehicles in use Europe 2018"
* "ACEA Report Vehicles in use Europe 2017"


```{r}
# View of 2019 car count data (untidied)
CarCount2019_untidy
# Cleaning the 2019 car count data
CarCount2019_untidy %>%
  extract(Data, c("Junk1", "Netherlands", "Junk2", "Norway", "Junk3"), 
          "(.*[Nethrlands]{11})([0-9,.-]+)(.*[Norway]{6})([0-9,.-]+)(.*)") %>%
  select(`Report Year`,`Vehicle Type`,`Netherlands`,`Norway`) %>% 
  pivot_longer(cols=c("Netherlands", "Norway"),names_to="Region",values_to="Counts") %>% 
  extract(Counts,c(2014:2018,"Junk"),
          "([0-9]{0,3}[,]?[0-9]{0,3},[0-9]{0,3})([0-9]{0,3}[,]?[0-9]{0,3},[0-9]{0,3})([0-9]{0,3}[,]?[0-9]{0,3},[0-9]{0,3})([0-9]{0,3}[,]?[0-9]{0,3},[0-9]{0,3})([0-9]{0,3}[,]?[0-9]{0,3},[0-9]{0,3})(.*)") %>% 
  select(-`Junk`) %>% 
  pivot_longer(cols=c("2014","2015","2016","2017","2018"),names_to="Year",values_to="Count1") %>%
  mutate(Count=as.numeric(gsub(",", "", `Count1`))) %>% 
  select(`Report Year`, `Region`, `Vehicle Type`, `Year`, `Count`) %>% 
  arrange(`Region`, `Vehicle Type`, desc(`Year`)) -> CarCount2019
CarCount2019
```

A display of the untidied and tidied 2019 car count data.


```{r}
# The 2018 and 2017 car count data is about the same.
clean_it <- function(x) {
  x %>%
    select(-contains("%")) %>% 
    pivot_longer(cols=contains("0"),names_to="Year",values_to="Count1") %>% 
    filter((Region=="Norway") | (Region=="Netherlands")) %>% 
    select(`Report Year`, `Region`, `Vehicle Type`, `Year`, `Count1`) %>% 
    arrange(`Region`, `Vehicle Type`, desc(`Year`)) %>% 
    mutate(Count=as.numeric(`Count1`)) %>% 
    select(-`Count1`) -> x_clean
  x_clean
}
(CarCount2018 <- clean_it(CarCount2018_untidy))
(CarCount2017 <- clean_it(CarCount2017_untidy))
```

A display of the tidied 2018 and 2017 car count data.


```{r}
# Merge the years from each report together where the latest report has the highest priority.
CarCount2019 %>% 
  rbind(CarCount2018, CarCount2017) %>% 
  unite(id,Region,`Vehicle Type`,Year,sep="_") %>%
  filter(!duplicated(id)) %>% # Keep data from latest report
  filter(!grepl("Netherlands_Total commercial vehicles \\(incl buses\\)",id)) %>% # Removed bc repeated data
  filter(!grepl(".*\\(incl buses\\)_201[2-5]",id)) %>% # Removed bc repeated data
  separate(id,into=c("Region","Vehicle Type","Year"),sep="_") %>% 
  arrange(`Region`, `Vehicle Type`, desc(`Year`)) -> CarCountALL
CarCountALL %>% unite(id,Region,`Vehicle Type`,Year,sep="_") -> CarCountALL_id
CarCountALL_id
```

In this dataset, the car count data from each report was merged. The latest records were kept for each year and repeated data was removed.


```{r}
# The fuel data is about the same for the 2018 and 2017 reports.
clean_it2 <- function(x) {
  x %>% 
    select(-Metric) %>% 
    filter((Region=="Norway") | (Region=="Netherlands")) %>%
    pivot_longer(Petrol:Other,names_to="Fuel Type",values_to="% Share") %>% 
    select(`Report Year`,`Region`,`Vehicle Type`, `Fuel Type`, `Year`, `% Share`) -> x_clean
  x_clean
}
FuelType2018.1 <- clean_it2(FuelType2018.1_untidy)
FuelType2018.2 <- clean_it2(FuelType2018.2_untidy)
FuelType2018.3 <- clean_it2(FuelType2018.3_untidy)
FuelType2017.1 <- clean_it2(FuelType2017.1_untidy)
FuelType2017.2 <- clean_it2(FuelType2017.2_untidy)
FuelTypeALL <- rbind(FuelType2018.1,FuelType2018.2,FuelType2018.3,FuelType2017.1,FuelType2017.2)
FuelTypeALL %>% unite(id,Region,`Vehicle Type`,Year,sep="_") -> FuelTypeALL_id
FuelTypeALL_id
```

In this dataset, the fuel data from the 2017 and 2018 report was merged.


```{r}
# Create a dataset containing the car count based on fuel type
FuelTypeALL_id %>%
  left_join(CarCountALL_id, by=c("id"="id")) %>% 
  mutate("Num Vehicles for Fuel Type"=`% Share`*`Count`) %>% 
  mutate("% Share for Fuel Type"=`% Share`) %>% 
  mutate("Total Vehicles (All Fuel Types)"=`Count`) %>% 
  separate(id,into=c("Region","Vehicle Type","Year"),sep="_") %>%
  select(`Region`,`Vehicle Type`, `Fuel Type`, `Year`, `% Share for Fuel Type`,
         `Num Vehicles for Fuel Type`, `Total Vehicles (All Fuel Types)`) %>% 
  arrange(`Region`,`Vehicle Type`, `Fuel Type`, `Year`) -> CarCountByFuel
CarCountByFuel
```

```{r}
# Export to excel file so it can be easily processed after
library(xlsx)
write.xlsx(CarCountByFuel, "NORWAY-NETHERcars.xlsx")
```
