---
title: "Data Prep: OECD Pollution"
---

```{r}
library(tidyverse)
pollutants_untidy <- read_csv("AirandGHGemissions.csv")
pollutants_untidy

# Check if we can remove the columns Indicator, Frequency, Flag Codes. There are 27120 rows in the original dataframe. So if all these have the same values, then we can remove the columns.
pollutants_untidy %>% count(INDICATOR=="AIREMISSION")
pollutants_untidy %>% count(FREQUENCY=="A")
pollutants_untidy %>% count(is.na(`Flag Codes`))
```
We can remove the 3 columns tested.


We will need to convert the 3 letter country codes into country names. So we need a file to convert it. 
```{r}
library(readxl)
codes <- read_excel("CountryCodes.xlsx",sheet="Data")
codes %>%
  separate(Country, c("Country","Junk"),sep="\\s\\(",remove=T) %>% 
  select(`Alpha-3 code`,Country) -> codes
codes$Country[codes$`Alpha-3 code` == "KOR"] <- "Republic of Korea"
codes
```


Based on the dashboard on the website (https://data.oecd.org/air/air-and-ghg-emissions.htm), we will use measure which offers the most data in terms of countries. These will be measures we'll use:
* CO2: Tonnes/capita
* CO: Kilograms/capita
* GHG: Tonnes/capita
* NOX: Kilograms/capita
* SOX: Kilograms/capita
* VOC: Kilograms/capita
Thus, we will filter only for "TONNE_CAP" and "KG_CAP".
```{r}
pollutants_untidy %>% 
  select(-c(INDICATOR,FREQUENCY,`Flag Codes`)) %>% 
  filter(MEASURE=="TONNE_CAP"|MEASURE=="KG_CAP") %>% 
  left_join(codes,by=c("LOCATION"="Alpha-3 code")) %>%
  mutate(id=paste(Country,TIME,sep="_"),Pollutant=SUBJECT,Measure=MEASURE,) %>% 
  select(id, Pollutant, Measure, Value) -> pollutants
pollutants
```


```{r}
# Export to excel file so it can be easily processed after
library(xlsx)
write.xlsx(pollutants, "pollutants.xlsx")
```


